---
title: "Rtricks"
author: "Vikesh Singh Baghel"
date: "23 February 2019"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today, I am going to show some examples on how to use data.table. R's data.table package is very fast and efficient. I will be using a simple data frame mtcars in this demo.

Let's load the data and see how what it has.

```{r }
data(mtcars)
head(mtcars)
```

To get the size of size, R has dim command. Let's also get summary statistics and class

```{r }
print(dim(mtcars))
print(summary(mtcars))
print(class(mtcars))
```

So, we have a data frame of 32 rows and 11 columns in this data set. All the columns are numeric.

I will store mtcars into a different object and start from there. Now I want to add another column to the data. I want to add avg_mpg_cg which will be mean of mpg over cyl and gear. There are several ways to do it. But my preferred approach is using data.table package.
```{r }
library(data.table)
dat <- as.data.table(mtcars)
dat[,avg_mpg_cg:= mean(mpg), by = c("cyl","gear")]
head(dat)
```


So, It can be in just one line! Okay, now may I dont need mpg column and want remove it.

```{r }
dat[,mpg:= NULL]
head(dat)
```

So, It can be in just one line! Okay, now may I dont need mpg column and want remove it.

```{r }
dat[,mpg:= NULL]
head(dat)
```
Boom! Done. Okay, that's cool. But wait mtcars had car names as rownames. It is not there in dat because data.table does not support rownames. But we can add a column to it.


So, It can be in just one line! Okay, now may I dont need mpg column and want remove it.

```{r }
dat[,car_names:= rownames(mtcars)]
head(dat)
```
Okay. So, now I want to select only the rows where car_names has Hornet

```{r }
dat[grep("Hornet",car_names)]
```

Hmmm. I want Mazda too.
```{r }
dat[grep(paste(c("Mazda","Hornet"),collapse="|"), car_names)]
```

Let's do now something more tricky. I want to apply a function (say mean) to disp, hp and drat column and store them in new columns of dat
```{r }
cols_to_work <- c("disp", "hp", "drat")
dat[,paste0(cols_to_work,"_avg"):= lapply(.SD, mean), .SDcols =cols_to_work ]
head(dat)
```




























